---
title: "Total Number of Rocket Launches By Country"
author: "M.Kee"
date: "`r format(Sys.time(), '%A, %B %d, %Y')`"
output:
  html_document:
    keep_md: no
    number_sections: no
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    highlight: pygments
    number_sections: yes
    toc: yes
    toc_depth: 1
  word_document:
    toc: yes
    toc_depth: '1'
---

## Project Overview
Humanity has always had an unyielding fascination with outer space. Over the past century in the United States, space operations have typically been planned, organized, and executed by government entities. With the decommissioning of the National Aeronautics and Space Administration (NASA) Space Shuttle program and the allowance and emergence for research and development of Commercial Orbital Transportation Services, new innovations are rapidly ushering in an unprecedented day of space operations and making it accessible to more people worldwide (Young & Kee, 2014)^[Young, J., & Kee, M. G. (2014). SpaceX Falcon 9/Dragon Operations NAS Impact and Operational Analysis. Atlantic City International Airport, NJ: US DOT Federal Aviation Administration.]. The goal of this data exploration exercise was to used the leaflet R package to visualize the total number of launches using active rockets by country and to visualize the total number of launches historically by country. 


## Data Description
The data used for this exercise are two Kaggle datasets, which are as follows: 

1. [Latitude and Longitude for Every Country & State](https://www.kaggle.com/paultimothymooney/latitude-and-longitude-for-every-country-and-state)

2. [Privatization of Space](https://www.kaggle.com/davidroberts13/one-small-step-for-data)


## Data Processing & Manipulation
```{r Initializing Workspace & Loading Data, message=FALSE, warning=FALSE, tidy=TRUE}
# Clearing variables before program execution
rm(list = ls())


# Loading necessary packages
library(tidyverse)
library(lubridate)
library(GGally)
library(RColorBrewer)
library(leaflet)
library(htmltools)
library(knitr)


# User-defined Variables that define desired working directory
cwdir <- "C:/Users/mkee1/Documents/Coursera-JH-Data-Science/09_Developing_Data_Products/Week 2/Peer Review Projects/Data/"


# Setting working directory using user-defined function "setdesiredwd"
source('~/setdesiredwd.R')
setdesiredwd(cwdir)


# Reading space launch & lat-long data into R using readr package
spaceDat <- read_csv("./GlobalSpaceLaunches/Global Space Launches.csv", col_types = cols(`Company Name` = "c", Location = "c", Detail = "c", `Status Rocket` = "c", Rocket = "n", `Status Mission` = "c", `Country of Launch` = "c", `Companys Country of Origin` = "c", `Private or State Run` = "f", DateTime = "_", Year = "_", Month = "_", Day = "_", Date = "_", Time = "_"))

spaceDat$`Country of Launch` <- recode(spaceDat$`Country of Launch`, "USA" = "United States")
spaceDat$`Companys Country of Origin` <- recode(spaceDat$`Companys Country of Origin`, "USA" = "United States")

names(spaceDat) <- c("Company Name", "Location", "Detail", "Status Rocket", "Rocket Cost (Millions)", "Status Mission", "Country of Launch", "Companys Country of Origin", "Private or State Run")

head(spaceDat)

LatLongWorld <- read_csv("./LatLong_WorldCountries/world_country_and_usa_states_latitude_and_longitude_values.csv", col_types = cols(country_code = "c", latitude = "d", longitude = "d", country = "c", usa_state_code = "_", usa_state_latitude = "_", usa_state_longitude = "_", usa_state = "_"))

LatLongWorld <- LatLongWorld %>% add_row(country_code = "PO", latitude = 0, longitude = -154, country = "Sea Launch")

names(LatLongWorld) <- c("Country Code", "Latitude (deg)", "Longitude (deg)", "Country")

head(LatLongWorld)
```


## Data Visualization: Total Number of Active Launches by Country
```{r Data Manipulation & Visualization 1, message=FALSE, warning=FALSE, tidy=TRUE}
# Filtering data to only include active rockets, summarizing based on Country of Launch & whether or not the launch is commercial/private or is for the State (or Government)
sumDat <- spaceDat %>% 
        filter(`Status Rocket` == "StatusActive") %>%
        select(`Country of Launch`, `Private or State Run`) %>%
        group_by(`Country of Launch`) %>%
        summarise(n())


# Joining the summary table & the lat-long data table together in order to get lat-long 
# information in the same table & renaming columns
jointData <- left_join(sumDat, LatLongWorld, by = c("Country of Launch" = "Country"))
names(jointData) <- c("Country of Launch", "Numuber Of Ops", "Country Code", "Latitude (deg)", "Longitude (deg)")
jointData <- jointData %>% arrange(desc(`Numuber Of Ops`))
kable(jointData, caption = "Total Number of Active Launches by Country")


# Code for data visualization using the leaflet R package. 
content <- paste0("<b>Country: </b>", jointData$`Country of Launch`, "<br/>",
                 "<b>Total Number of Launches: </b>", jointData$`Numuber Of Ops`)


leaflet(jointData) %>% addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addAwesomeMarkers(~`Longitude (deg)`, ~`Latitude (deg)`, popup = content)
```


## Data Visualization: Total Number of Launches Historically by Country
```{r Data Manipulation & Visualization 2, message=FALSE, warning=FALSE, tidy=TRUE}
# Summarizing based on Country of Launch & whether or not the launch is commercial/private or is  for the State (or Government)
sumDat2 <- spaceDat %>% 
        select(`Country of Launch`, `Private or State Run`) %>%
        group_by(`Country of Launch`) %>%
        summarise(n())


# Joining the summary table & the lat-long data table together in order to get lat-long 
# information in the same table & renaming columns
jointData2 <- left_join(sumDat2, LatLongWorld, by = c("Country of Launch" = "Country"))
names(jointData2) <- c("Country of Launch", "Numuber Of Ops", "Country Code", "Latitude (deg)", "Longitude (deg)")
jointData2 <- jointData2 %>% arrange(desc(`Numuber Of Ops`))
kable(jointData2, caption = "Total Number of Launches Historically by Country")

# Code for data visualization using the leaflet R package. 
content <- paste0("<b>Country: </b>", jointData2$`Country of Launch`, "<br/>",
                 "<b>Total Number of Launches: </b>", jointData2$`Numuber Of Ops`)


leaflet(jointData2) %>% addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addAwesomeMarkers(~`Longitude (deg)`, ~`Latitude (deg)`, popup = content)
```